---
title: "U.S.A Incomes by Occupation and Gender "
subtitle: "by Roey Nir and Tamar Tal"
date: "24.6.2021"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



<br>
<br>

#### *introduction*
In this paper we will analyze the ['Ask A Manager Salary Survey 2021']( https://docs.google.com/spreadsheets/d/1IPS5dBSGtwYVbjsfbaMCYIWnOuRmJcbequohNxCyGVw/edit?resourcekey#gid=1625408792) data set. 

This data set contains information regarding people's income and several different variable as you will see below.     
We would like to examine the effects of different variables such as occupation, gender and years of experience on one's annual salary in three major fields: Hi-tech, Health care and Higher education.

***

#### *Data import and tidying*

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(tidyr)
library(tidyverse)
library(tidyr)
library(datasets)
library(ggplot2)
library(readxl)
library(janitor)
library(data.table)
library(stringi)
library(unglue)
```

```{r importing_the_data, echo=FALSE,message=FALSE,warning=FALSE }
data<- read_excel("Ask A Manager Salary Survey 2021.xlsx")
```

First, we need to adjust the data set - here are the main adjustments we did: 

* We filtered the data so we can work with information regarding the USA and the relevant occupation.  
* We selected the relevant columns for our hypothesis and questions.  
* We deleted 'NA' values, unclassified gender and unreasonable salaries. 

```{r cleaning, echo=FALSE,message=FALSE,warning=FALSE }
us_employ <- data %>%
  filter(Currency=="USD", Gender =="Man"|Gender=="Woman", `Annual salary`!=0 , `Annual salary`!= 102000000, `Annual salary`!= 1)

us_employ <- us_employ %>%
  select(`How old are you?`,Industry,`Annual salary`,`Overall years of professional experience`,`Highest level of education completed`,Gender,Race)

us_employ<-us_employ%>% drop_na()

us_employ<-us_employ%>% clean_names()

us_employ<- us_employ%>%
  filter(industry=="Computing or Tech"|industry=="Health care"|industry=="Education (Higher Education)")
```

#### *Transformation*

We did a little bit of transformation so it will be easier for us to work with the data and for you to go over it – here are the main changes we did:

* We divided the 'Annual salary' column so that we will work with smaller numbers.    
* We changed the 'highest level of education completed' values to degrees name that matches the types of degrees in Israel.    
* We changed the columns names according to the transformations above.

<br>

Here, take a little glimpse…

```{r transformations, echo=FALSE,message=FALSE,warning=FALSE}
us_employ<- us_employ%>%
  mutate(annual_salary=annual_salary/1000)

us_employ[us_employ == "Education (Higher Education)"]<-"Higher Education"
us_employ[us_employ == "Computing or Tech"]<-"Hi_tech"

us_employ$highest_level_of_education_completed[us_employ$highest_level_of_education_completed %like% "llege"] <- "Bachelor Degree"
us_employ$highest_level_of_education_completed[us_employ$highest_level_of_education_completed %like% "Mas"] <- "Master's Degree"
us_employ$highest_level_of_education_completed[us_employ$highest_level_of_education_completed %like% "Ph"|us_employ$highest_level_of_education_completed %like% "rofessio"] <- "Doctoral Degree"
us_employ$highest_level_of_education_completed[us_employ$highest_level_of_education_completed %like% "High"] <- "High School Graduate"

glimpse(us_employ)
```

***
#### *Visualizations & Modeling*

The next plots will show the difference between man's and woman's average annual salary in each field:

<br>
<br>

```{r first_plots, echo=FALSE, message=FALSE,warning=FALSE,out.width = "50%",fig.show='hold'}

plot1<-us_employ%>%
  group_by(industry,gender)%>%
  summarise(average_salary=mean(annual_salary))

ggplot(aes(industry, average_salary,fill=gender),data=plot1)+
  geom_col(position = 'dodge')+
  scale_fill_manual(values = c("#7cb6fd","#fdc27c"))+
  labs(title = "Average Annual Salary",y="Average salary (in thousands)",x="Field")+
  geom_text(aes(label=round(average_salary,1)),color="white",vjust=1.5,
            position = position_dodge(0.9))+
  theme(panel.background = element_rect(fill = "white"), legend.position = c("0.90","0.92"))




m<- us_employ %>%
  filter(gender=="Man")%>% select(annual_salary)
m_mean<- mean(m$annual_salary, na.rm = TRUE)

woman<- us_employ%>%
  filter(gender=="Woman")%>% count()

over<- us_employ%>%
  filter(gender=="Woman",annual_salary>=m_mean)%>% count()/4913 *100
less<- us_employ%>%
  filter(gender=="Woman",annual_salary<m_mean)%>% count()/4913 *100


df<- data.frame(group=c("Earn more than an average man's salary","Earn less than an average man's salary"),value=c(15.26562,84.73438))


df%>%
  ggplot(aes(x="",y=value,fill=group))+
  geom_bar(width = 1,stat="identity",color="white")+
  coord_polar("y",start=0)+
  geom_text(aes(label= paste0(round(value),"%")),position = position_stack(vjust = 0.5) ,color="black")+
  scale_fill_manual(values = c("#feb956","#fee2b8"))+
  labs(x= NULL, y=NULL, fill=NULL, title = "Women's income compared to men's average income")+
  theme_classic()+
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = c("0.99","0.92"))


```

```{r pai chart ,echo=FALSE,message=FALSE,warning=FALSE,out.width = "50%", out.extra='style="float:right;padding:10px"'}

plot2 <- us_employ %>%
  select(industry,gender,annual_salary)

plot2%>%
  ggplot(aes(x=gender,y=annual_salary, color=gender))+
  geom_boxplot(alpha=0.3,outlier.shape=NA, width=1,color=c("#7cb6fd","#fdc27c","#7cb6fd","#fdc27c","#7cb6fd","#fdc27c"),
               fill=c("#7cb6fd","#fdc27c","#7cb6fd","#fdc27c","#7cb6fd","#fdc27c"))+
  facet_wrap(~industry)+
  theme(legend.position = "none",panel.border = element_blank())+
    ylim(0,200)+
  labs(x= "Gender",
       y="Annual salary (in thousands)",
       title="Annual salary median")+
  theme_light()
```

<br>


As we can see in these charts, in every field, woman's income is much lower.  
We can also see the low percentage of woman whose salary is equal or higher than the average salary of man.

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

***

##### *First model*
Clearly gender has an effect on one's income.
Now we would like to use unpaired two samples t-test to show our hypothesis with a significance level of 0.05.

*Our hypothesis:*  
\(H_0\)= there is no income difference between genders.  
\(H_1\)= man's income is higher than woman's income.

The samples are unpaired, so we first need to use Levene`s test to check if the variance of both samples are equal.

Levene`s test:
```{r t-test,echo=FALSE,message=FALSE,warning=FALSE}
t_test_data<-us_employ%>%select(gender,annual_salary)%>%drop_na()
```

```{r f-test,echo=FALSE,message=FALSE,warning=FALSE}
var.test(t_test_data$annual_salary~ t_test_data$gender, alternative = "two.sided")
```

As we can see, the p-value equals to 2.2e-16 and smaller than 0.05, therefore, we **can reject** the \(H_0\) assumption, which means the variances are different.

Let`s move on to the one tailed t-test:
```{r one-tailed-t-test,echo=FALSE,message=FALSE,warning=FALSE}
t.test(t_test_data$annual_salary ~ t_test_data$gender, var.equal = F, alternative = "greater")
```
The p-value equals to 2.2e-16 and smaller than 0.05, therefore, we **can reject** the \(H_0\) assumption, which means that men income is higher than women income. 

***

#### *More visualizations...*

The next question we want to discuss, is if there is a linear connection between one`s total years of experience and his annual income.  


```{r import_reg_lin,echo=FALSE,message=FALSE,warning=FALSE}
lin_reg <- us_employ %>% select(industry, annual_salary, overall_years_of_professional_experience)%>% group_by(industry) %>% slice(501:550)
lin_reg$overall_years_of_professional_experience <- gsub('[-]', ':', gsub('[a-zA-Z ]', '', lin_reg$overall_years_of_professional_experience))

lin_reg$overall_years_of_professional_experience[lin_reg$overall_years_of_professional_experience=="1"] <- 0
lin_reg$overall_years_of_professional_experience[lin_reg$overall_years_of_professional_experience=="2:4"] <- 3
lin_reg$overall_years_of_professional_experience[lin_reg$overall_years_of_professional_experience=="5:7"] <- 6
lin_reg$overall_years_of_professional_experience[lin_reg$overall_years_of_professional_experience=="8:10"] <- 9
lin_reg$overall_years_of_professional_experience[lin_reg$overall_years_of_professional_experience=="11:20"] <- 15
lin_reg$overall_years_of_professional_experience[lin_reg$overall_years_of_professional_experience=="21:30"] <- 25
lin_reg$overall_years_of_professional_experience[lin_reg$overall_years_of_professional_experience=="31:40"] <- 35
lin_reg$overall_years_of_professional_experience[lin_reg$overall_years_of_professional_experience=="41"] <- 45


lin_reg$overall_years_of_professional_experience <- as.numeric(lin_reg$overall_years_of_professional_experience)
```


```{r second_plot,echo=FALSE,message=FALSE,warning=FALSE,out.width = "65%", out.extra='style="float:right;padding:10px"'}
lin_reg %>% ggplot(aes(x = overall_years_of_professional_experience, y = annual_salary, color=annual_salary)) +
  geom_point()+
  scale_colour_gradient(low ="#0675fb",high="#f42a86" )+
  facet_wrap(~industry)+
  scale_x_continuous(breaks = seq(5,50,by=10))+
  scale_y_continuous(breaks = seq(0,350,by=50))+
  labs(x= "Years of experience",
       y="Annual salary (in thousands)",
       title="Annual salary distribution depending on years of experience")+
    theme(legend.title = element_blank())

```
<br>
<br>

For this part, we created a new variable with the necessary columns and changed the 'overall years of professional experience' column to the middle value of the given range.     
Next we used the function 'sample_n' to sample 50 random rows from each industry, so that way we can work with a database big enough but not too big.

<br>
<br>

At first, we thought it would be easy to see a correlation between years of experience and salary but while looking at the plot, we see that this is not the case; there are plenty of people with many years of experience and lower salaries.

We will now try to test this with linear regression model.

<br>
<br>

***

##### *second model*
We are using linear regression model with significance level of 0.05.
Of course the dependent variable $Y$ is the annual salary and the independent variable $X$ is the number of years of experience.


*Our hypothesis:*  
\(H_0\)= there in no linear connection between $X$ and $Y$.  
\(H_1\)= The larger the $X$, the larger the $Y$.

Another way to write the hypothesis is:  
\(H_0\): $\beta_1$=0  
\(H_1\): $\beta_1$>0


We chose to examine each field individually because the salary distribution between the fields is different:

1. Health care:
```{r reg_health_care,echo=FALSE,message=FALSE,warning=FALSE}
health <- lin_reg %>% filter(industry == "Health care")
lm2 <- lm(annual_salary ~ overall_years_of_professional_experience, data = health)
summary(lm2)

```
<br>

As we can see, the p-value equals to 0.5637 and bigger than 0.05, therefore, we **cannot reject** the \(H_0\) assumption.  
The conclusion is that there is no linear connection between years of experience and annual salary in Healthcare field.

***

2. Hi-tech:
```{r reg_hi_tech,echo=FALSE,message=FALSE,warning=FALSE}
tech <- lin_reg %>% filter(industry == "Hi_tech")
lm1 <- lm(annual_salary ~ overall_years_of_professional_experience, data = tech)
summary(lm1)

```
<br>

Here the p-value equals to 0.02637 and smaller than 0.05, therefore, we **can reject** the \(H_0\) assumption.  
In contrast to the last field there is a positive linear connection between years of experience and annual salary in Hi-tech field.    

***

3. Higher education:
```{r reg_education,echo=FALSE,message=FALSE,warning=FALSE}
education <- lin_reg %>% filter(industry == "Higher Education")
lm3 <- lm(annual_salary ~ overall_years_of_professional_experience, data =education )
summary(lm3)
```
<br>

Again, there is a positive linear connection between the number of years of experience and the annual salary.    
We can see that the p- value equals to 0.00007 and smaller than 0.05.

***
```{r lin_reg_plot,echo=FALSE,message=FALSE,warning=FALSE,out.width = "65%", out.extra='style="float:right;padding:10px"'}
ggplot(lin_reg) +
  geom_point(aes(x = overall_years_of_professional_experience, y = annual_salary,color=annual_salary)) + 
  geom_smooth(method=lm, aes(x = overall_years_of_professional_experience, y = annual_salary),colour="#55595f",size=0.8) + 
  facet_wrap(~industry)+
  scale_colour_gradient(low ="#0675fb",high="#f42a86" )+
  scale_x_continuous(breaks = seq(5,50,by=10))+
  scale_y_continuous(breaks = seq(0,350,by=50))+
  labs(x= "Years of experience",
       y="Annual salary (in thousands)",
       title="Linear regression of annual salary depending on years of experience")+
  theme(legend.title = element_blank())
```
<br>

Let's look at the last field - Hi-tech...  
We rejected the \(H_0\) assumption and determined that there is a linear connection.  
However, how strong is that connection? For that, we can look at the R-squared value which is the indicator for the goodness of the fit between our two variables. In this case, R-squared is low so it can be said that our model does not explain well the relationship between the variables.  
Maybe next time we would prefer to do other tests or use a larger sample.

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

***

 
#### *Summary*

We examined a couple of question regarding income, gender and occupation.  
We saw the wage gap between genders through plots and a t-test and we got an general idea about the meaning of experience in the market using regression analysis.  
In addition, this project gave us a chance to practice different methods in R programming and also some experience that will help us in the future.   

Thank you!

<br>
<br>
